A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-07-24 paper/elsarticle/elsarticle-template.tex
paper

Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie. Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}; the calibration matrix~$\mathtt{K}$ is then
\begin{equation}
\mathtt{K} = \begin{bmatrix}
f & 0 & x_0 \\
0 & f & y_0 \\
0 & 0 & 1
\end{bmatrix},
\end{equation}
where $f$ and $(p_x, p_y)^\top$ are both expressed in units of pixels.

In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}. To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is used to obtain the lens distortion model coefficients, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
\begin{equation}
\sum_{i,j} d\big(\mathbf{x}_{i,j}, \mathtt{K}(\mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j)\big),
\label{re}
\end{equation}
where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$ calibration image, $(\mathbf{x}_{i,j}', 1)^\top \sim \mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{P}^2$ gives the projection~$\mathbf{x}'_{i,j} \in \mathbb{R}^2$ of $\mathbf{X}_{i,j}'$ to that image, and $d$ is a distance function that~(i) corrects for lens distortion using the recovered lens distortion model coefficients~\cite{duane1971close} and (ii) returns a distance with respect to the two undistorted pixel positions. The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.

@@

\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed. 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.

>>>>

Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which the projecting point~$\mathbf{X}$ must lie. Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii)~the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close,weng1992camera}; the calibration matrix~$\mathtt{K}$ is then
\begin{equation}
\mathtt{K} = \begin{bmatrix}
f & 0 & x_0 \\
0 & f & y_0 \\
0 & 0 & 1
\end{bmatrix},
\end{equation}
where $f$ and $(p_x, p_y)^\top$ are both expressed in units of pixels.

In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}. To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is used to obtain the lens distortion model coefficients, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
\begin{equation}
\sum_{i,j} d\big(\mathbf{x}_{i,j}, \mathtt{K}(\mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j)\big),
\label{re}
\end{equation}
where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$ calibration image, $(\mathbf{x}_{i,j}', 1)^\top \sim \mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{P}^2$ gives the projection~$\mathbf{x}'_{i,j} \in \mathbb{R}^2$ of $\mathbf{X}_{i,j}'$ to that image, and $d$ is a distance function that~(i) corrects for lens distortion using the lens distortion model coefficients~\cite{duane1971close,weng1992camera} and (ii) returns a distance with respect to the two undistorted pixel positions. The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.

@@

\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed. 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.
